---
title: "Class 9"
author: "Peng Fei Huang"
date: "5/2/2018"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Unsupervised Learning Analysis of Cancer Cells

Lets read our input data 

```{r}
url <- "https://bioboot.github.io/bggn213_S18/class-material/WisconsinCancer.csv"

# Complete the following code to input the data and store as wisc.df
wisc.df <- read.csv(url)
head(wisc.df)
```

How many diagnosis are cancer vs non cancer 
```{r}
table(wisc.df$diagnosis)
```

Lets make a new data matrix with just the numeric values of interst in it. So we want to get rid of the first 2 columns 
```{r}
# Convert the features of the data: wisc.data
# Use subscript - to delete stuff
#x <- c("barry", "chris", "mary", "pete")
#x[-c(1,4)]
#as.numeric(x == "mary")
## Also remove the problem 'X' column in position 33 
wisc.data <- as.matrix( wisc.df[,-c(1:2,33)] )
rownames(wisc.data) <- wisc.df$id
diagnosis <- as.numeric(wisc.df$diagnosis == "M")
sum(diagnosis)
```


Q1. How many observations are in this dataset?
569 or could use dim() whi
```{r}
nrow(wisc.data)

```

Q2. How many variables/features in the data are suffixed with _mean?
31
```{r}
# number of columns
ncol(wisc.data)
# names of all the columns
colnames(wisc.data)
#grabing the specific name 
#grep("_mean", colnames(wisc.data), value=TRUE, invert = TRUE)
grep("_mean", colnames(wisc.data), value=TRUE)
length(grep("_mean", colnames(wisc.data), value=TRUE))
```

Q3. How many of the observations have a malignant diagnosis?
212
```{r}
table(wisc.df$diagnosis)
```


## Principal Component Analysis 
Check the mean and standard deviation of the features (i.e. columns) of the wisc.data to determine if the data should be scaled. Use the  colMeans() and apply() functions like you’ve done before.

```{r}
# Check column means and standard deviations
plot(colMeans(wisc.data), type="o")
colMeans(wisc.data)
apply(wisc.data,2,sd)
```


```{r}
# Perform PCA on wisc.data by completing the following code
wisc.pr <- prcomp(wisc.data, scale = TRUE )
```

How well are we doing...
```{r}
summary(wisc.pr)
```

## PCA results plot

```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2], col = diagnosis+1)
#Red is cancer, Black is noncancer

```


## Scree-plot: variance explained 

```{r}
# Variance explained by each principal component: pve
pve <- wisc.pr$sdev^2 / sum(wisc.pr$sdev^2)

# Plot variance explained for each principal component
plot(pve, xlab = "Principal Component", 
     ylab = "Proportion of Variance Explained", 
     ylim = c(0, 1), type = "o")
```


Lets do a barplot of this data  
```{r}
barplot(pve, names.arg=paste("PC", 1:length(pve)), las=2, axe = FALSE, ylab = "Proportion of Variance")
axis( 2, at=pve, labels = round(pve,2)*100)
```


## Hierarchical clustering 

First we want to scale our data using scale() function

```{r}
# Scale the wisc.data data: data.scaled
data.scaled <- scale(wisc.data)
```

#Calculate the (Euclidean) distances between all pairs of observations in the new scaled dataset and assign the result to data.dist.

```{r}
data.dist <- dist(data.scaled)

```
#Create a hierarchical clustering model using complete linkage. Manually specify the method argument to hclust() and assign the results to  wisc.hclust.

```{r}
wisc.hclust <- hclust(data.dist, method = "complete" )

```
# Lets plot our tree 
```{r}
plot(wisc.hclust)
abline(h=20, col="red", lwd=3)
```



```{r}
wisc.hclust.clusters <- cutree(wisc.hclust, k=4)

```
### how do these groups match our diagnosis 
```{r}
table(diagnosis)
table(wisc.hclust.clusters)
```
```{r}
table(wisc.hclust.clusters, diagnosis)
```



Input our scaled values here
#No idea why my stuff is in inverse
```{r}
data.scaled <- scale(wisc.data)
wisc.km <- kmeans(data.scaled, centers= 2, nstart= 20)
```

Compare to expert diagnosis
```{r}
table(wisc.km$cluster)
table(wisc.km$cluster, diagnosis)
```

```{r}
table(wisc.hclust.clusters, wisc.km$cluster)
```



```{r}
## Use the distance along the first 7 PCs for clustering i.e. wisc.pr$x[, 1:7]
wisc.pr.hclust <- hclust(dist(wisc.pr$x[, 1:3]), method="ward.D2")
plot(wisc.pr.hclust)
```

```{r}
wisc.pr.hclust.clusters <- cutree(wisc.pr.hclust, k=2)
plot(wisc.pr$x[, 1:2], col=wisc.hclust.clusters)
```

```{r}
table(wisc.pr.hclust.clusters, diagnosis)

```



```{r}
##install.packages("rgl")
#library(rgl)
```

##Bonus

```{r}
url <- "https://tinyurl.com/new-samples-CSV"
new <- read.csv(url)
npc <- predict( wisc.pr, newdata=new)
npc
plot(wisc.pr$x[,1:2], col=wisc.pr.hclust.clusters)
points(npc[,1], npc[,2], col=c("purple","blue"), pch=16, cex=3)
```






















