---
title: "Class 14"
author: "Peng Fei Huang"
date: "5/18/2018"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Transcriptiomics and the analysis of RNA-Seq data 

Input our count matrix and meta-data matrix about our experiment setup

```{r}
counts <- read.csv("data/airway_scaledcounts.csv", stringsAsFactors = FALSE)
metadata <-  read.csv("data/airway_metadata.csv", stringsAsFactors = FALSE)
```

```{r}
colnames(counts)[-1] == metadata$id
```
Calculate the mean counts for each gene across all control experiment
```{r}
control <- metadata[metadata[,"dex"]=="control",]
control
control.mean <- rowSums( counts[ ,control$id] )/nrow(control) 
names(control.mean) <- counts$ensgene
```


```{r}
head(control.mean)
```

Now calculate the mean counts for each gene across all drug experiment
```{r}
treated <- metadata[metadata[,"dex"]=="treated",]
treated
treated.mean <- rowSums( counts[ ,treated$id] )/nrow(treated) 
names(treated.mean) <- counts$ensgene
```

Combine into a new data.frame called 'meancounts'
```{r}
meancounts <- data.frame(control.mean, treated.mean)
head(meancounts)
```

```{r}
colSums(meancounts)

```


Plot our mean counts  for control vs. trated
```{r}
plot(meancounts$control, meancounts$treated)
```

Plot with log exists so we see our data 
```{r}
plot.default(meancounts$control, meancounts$treated)
```


Here we calculate log2foldchange, add it to our meancounts data.frame
```{r}
meancounts$log2fc <- log2(meancounts[,"treated.mean"]/meancounts[,"control.mean"])
head(meancounts)
```

Check our understanding of the ** which() ** function
```{r}
x <- matrix(1:10, ncol=2, byrow= TRUE)
x
x[5.2] <-0
x==0
which(x==0, arr.ind = TRUE)


which(c(T,F,T))
```


Now remove the zero count containing geens 
```{r}
zero.vals <- which(meancounts[,1:2]==0, arr.ind=TRUE)

to.rm <- unique(zero.vals[,1])
mycounts <- meancounts[-to.rm,]
head(mycounts)
```

```{r}
up.ind <- mycounts$log2fc > 2
down.ind <- mycounts$log2fc < (-2)
```


```{r}
head(up.ind)
paste("Up:", sum(up.ind))
paste("Down", sum(down.ind))
```

Here we Add annotation from a supplied CSV file, such as those available from ENSEMBLE or UCSC. 
```{r}
anno <- read.csv("data/annotables_grch38.csv")
head(anno)
```

Use the *merge* function
```{r}
results <- merge(mycounts, anno, by.x="row.names", by.y="ensgene")
head(results)
```

##another way to annotate my gene lim
```{r}
#biocLite("org.Hs.eg.db")
#biocLite("AnnotationDbi")
library("AnnotationDbi")
library("org.Hs.eg.db")
```


```{r}
mycounts$symbol <- mapIds(org.Hs.eg.db,
                     keys=row.names(mycounts),
                     column="SYMBOL",
                     keytype="ENSEMBL",
                     multiVals="first")
```


# DESeq2 Analysis 

Setup our DESeq Data set a 
```{r}
library("DESeq2")
dds <- DESeqDataSetFromMatrix(countData=counts, 
                              colData=metadata, 
                              design=~dex, 
                              tidy=TRUE)
dds
```

```{r}
#results(dds)
```

```{r}
dds <- DESeq(dds)
```
We can get results out of the object simply by calling the results() function on the DESeqDataSet that has been run through the pipeline.
```{r}
res <- results(dds)
res
```



We can summarize some basic tallies using the summary function 
```{r}
summary(res)
```

Order our results by p-values 
```{r}
resOrdered <- res[order(res$pvalue),]
head(resOrdered)
```


```{r}
res05 <- results(dds, alpha=0.05)
summary(res05)
```

#Q9. How many are significant with an adjusted p-value < 0.05? How about 0.01? Save this last set of results as resSig01.
```{r}
resSig05 <- subset(as.data.frame(res), padj < 0.05)
nrow(resSig05)
```

```{r}
resSig01 <- subset(as.data.frame(res), padj < 0.01)
nrow(resSig01)
```

#Q10. Using either the previously generated anno object (annotations from the file annotables_grch38.csv file) or the mapIds() function (from the AnnotationDbi package) add annotation to your res01 results data.frame.


```{r}
resSig01$symbol <- mapIds(org.Hs.eg.db,
                     keys=row.names(resSig01),
                     column="SYMBOL",
                     keytype="ENSEMBL",
                     multiVals="first")
head(resSig01)
```


Order by adjusted pvalue or log2fold change and write out to a CSV file
```{r}
ord <- order( resSig01$padj )
#View(res01[ord,])
head(resSig01[ord,])
```

```{r}
write.csv(resSig01[ord,], "signif01_results.csv")
```

#Data Visualization
```{r}
i <- grep("CRISPLD2", resSig01$symbol)
resSig01[i,]
rownames(resSig01[i,])

```


```{r}
plotCounts(dds, gene="ENSG00000103196", intgroup="dex")

```


```{r}
# Return the data
d <- plotCounts(dds, gene="ENSG00000103196", intgroup="dex", returnData=TRUE)
head(d)
```

```{r}
boxplot(count ~ dex , data=d)

```

```{r}
library(ggplot2)
ggplot(d, aes(dex, count)) + geom_boxplot(aes(fill=dex)) + scale_y_log10() + ggtitle("CRISPLD2")
```


#MA & Volcano plots

```{r}
res$sig <- res$padj<0.05

# How many of each?
table(res$sig)
```

```{r}
sum(is.na(res$sig))

```

#In Built MA plot
```{r}
plotMA(res, ylim=c(-2,2))

```


```{r}
resLFC <- lfcShrink(dds, coef=2)
resLFC
```


```{r}
plotMA(resLFC, ylim=c(-2,2))

```


```{r}
ggplot(as.data.frame(res), aes(log2FoldChange, -1*log10(pvalue), col=sig)) + 
    geom_point() + 
    ggtitle("Volcano plot")
```


#Side-note: Transformation
```{r}
vsdata <- vst(dds, blind=FALSE)

```

#PCA
```{r}
plotPCA(vsdata, intgroup="dex")

```

Session Information
```{r}
sessionInfo()

```











