---
title: "Bioinformatics Lecture 7"
author: "Peng Fei Huang"
date: "4/25/2018"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Functions again

We can source any file of R code with the 'source' () function

```{r}
source("http://tinyurl.com/rescale-R") 
```

Lets make sure things are here 

```{r}
ls()
```

Check out rescale() is working 
```{r}
rescale(1:10)
```

```{r, eval=FALSE}
rescale(c(1:10, "string"))
```

Let's check if rescale2 does any better 

```{r,eval=FALSE}
rescale2(c(1:10, "strings"))
```



## Function for finding missing values in two datasets

Write a both_na() function to do this 

```{r}
x <- c( 1, 2, NA, 3, NA)
y <- c(NA, 3, NA, 3, 4)
is.na(x)
```

```{r}
which(is.na(x))
```

```{r}
sum(is.na(x))
```
 
```{r}
is.na(x)
is.na(y)
```
We can use AND here 
```{r}
is.na(x) & is.na(y)
```

How many TRUE things are in this vector
```{r}
sum(is.na(x) & is.na(y))
```

My first function can start from this snippet 
```{r}
both_na <- function(x,y) {
  sum(is.na(x) & is.na(y))
}
```
Test is 
```{r}
both_na(x,y)
```

Testing 

```{r}
x <- c(NA, NA, NA)
y1 <- c( 1, NA, NA)
y2 <- c( 1, NA, NA, NA)

both_na(x, y2)
```


```{r}
#both_na2(x,y2)
```



```{r}
x <- c( 1, 2, NA, 3, NA)
y <- c(NA, 3, NA, 3, 4)

ans <- both_na3(x,y)
```

```{r}
ans$number
ans$which 
```



## And a last funciton tha tis actually useful 

```{r}
x <-  df1$IDs
y <- df2$IDs

x
y
```

We can try the intersect function and the %in% function 
```{r}
intersect(x,y)
x %in% y 
```


We can use the logical output in %in% to get at our matching data ... 
```{r}
x[x %in% y]
y[y %in% x]
```


Lets put these together as columns of a matrix
```{r}
cbind(x[x %in% y], y[y %in% x])
```


```{r}
cbind(c("Hello", "Help"), c("Please", "Me"))
rbind(c("Hello", "Help"), c("Please", "Me"))

```

Now we can make our first function
Test it on x and y
```{r}
function(x, y) { 
   cbind( x[ x %in% y ], y[ y %in% x ] )
}

gene_intersect(x,y)
```


Lets try with data.frame input rather than vectors 

```{r}
gene_intersect2(df1, df2)
```

Looks good, this is our skateboard 

Add some flexibility for col name to match by 

```{r}
gene_intersect3(df1, df2)
```



Lets use the merge function for this 
```{r}
merge(df1, df2, by="IDs")
```





