---
title: "Bioinformatics Class 11"
author: "Peng Fei Huang"
date: "5/9/2018"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## PDB Statistics 

Import out PDB statistics CSV file and calculate precent strucutres by experimental method 
```{r}
p <- read.csv("Data Export Summary.csv", row.names = 1)
```


```{r}
percent <- (p$Total/sum(p$Total)) *100
names(percent) <-  row.names(p)
percent

```


## Using Bio3D

Load the bio3d package


```{r}
library(bio3d)
```

Read in our HIV-Pr structure 

```{r}
pdb <- read.pdb("1hsg")
pdb
```

```{r}
attributes(pdb)
```

```{r}
head(pdb$atom)
```

# Print a subset of $atom data for the first two atoms
```{r}
pdb$atom[1:2, c("eleno", "elety", "x","y","z")]
```

# Note that individual $atom records can also be accessed like this
```{r}
pdb$atom$elety[1:2]
```

# Which allows us to do the following
```{r}
#plot.bio3d(pdb$atom$b[pdb$calpha], sse=pdb, typ="l", ylab=“B-factor”)
```

# Print a summary of the coordinate data in $xyz
```{r}
pdb$xyz
```

# Examine the row and column dimensions
```{r}
dim(pdb$xyz)
```

# Print coordinates for the first two atom
```{r}
pdb$xyz[ 1, atom2xyz(1:2) ]
```

# Select all C-alpha atoms (return their indices)
```{r}
ca.inds <- atom.select(pdb, "calpha")
ca.inds
```


# Print details of the first few selected atoms
```{r}
head( pdb$atom[ca.inds$atom, ] )
```


# And selected xyz coordinates
```{r}
head( pdb$xyz[, ca.inds$xyz] )
```


# Select chain A
```{r}
a.inds <- atom.select(pdb, chain="A")
```

# Select C-alphas of chain A
```{r}
ca.inds <- atom.select(pdb, "calpha", chain="A")
```

# We can combine multiple selection criteria to return their intersection
```{r}
cab.inds <- atom.select(pdb, elety=c("CA","CB"), chain="A", resno=10:20)
```




#Q8. Use the Bio3D write.pdb() function to write out a protein only PDB file for viewing in VMD. Also write out a second separate PDB file for the ligand with residue name MK1

```{r}
inds.ligand <- atom.select(pdb, "ligand")
inds.protein <- atom.select(pdb, "protein")
inds.protein
```


Check we have what we want
```{r}
pdb$atom[inds.ligand$atom,]
```

```{r}
head(pdb$atom[inds.protein$atom,])

```

```{r}
pdb.ligand <- trim.pdb(pdb, inds=inds.ligand)
pdb.ligand
pdb.protein <-  trim.pdb(pdb,inds=inds.protein)
```


```{r}
write.pdb(pdb.ligand, file="1hsg_ligand.pdb")
write.pdb(pdb.protein, file="1hsg_protein.pdb")
```


## Download some example PDB files
## Aligning mutliple strucutres
```{r}
ids <- c("1TND_B","1AGR_A","1TAG_A","1GG2_A","1KJY_A","4G5Q_A")
files <- get.pdb(ids, split = TRUE)
```

# Extract and align the chains we are interested in
```{r}
pdbs <- pdbaln(files, fit = TRUE)
pdbs
```


##Section 5: Example Application on Adenylate Kinase (Adk)

```{r}
aa <- get.seq("1ake_A")
```

# Blast or hmmer search 
```{r}
b <- blast.pdb(aa)
```

# Plot a summary of search results
```{r}
hits <- plot(b)
hits
```

```{r}
head(hits$pdb.id)
```

# Fetch PDBs
```{r}
files <- get.pdb(hits$pdb.id, path = "pdbs", split = TRUE, gzip = TRUE)
```

# Align structures
```{r}
pdbs <- pdbaln(files)
```

# Vector containing PDB codes
```{r}
ids <- basename.pdb(pdbs$id)
```

# Draw schematic alignment
```{r}
plot(pdbs, labels=ids)
```


##Principle Component Analysis
```{r}
# find invariant core
core <- core.find(pdbs)

```


```{r}
# superimpose all structures to core
pdbs$xyz = pdbfit(pdbs, core)
```


```{r}
# Perform PCA
pc.xray <- pca(pdbs)
plot(pc.xray)
```


# Visualize first principal component
```{r}
pc1 <- mktrj(pc.xray, pc=1, file="pc_1.pdb")
```























